---
title: "第2回: Mixed model"
output: html_notebook
date: 2018-04-23 16:00--
---

## Mixed model

[Environmental computing の mixed-model のページ](http://environmentalcomputing.net/mixed-models/)

* random variable

4 criteria

* is categorical
* has a large number of levels
* only a random subsample of levels is included in your design
* you want to make inference in general, and not only for the levels you observed.


薬学とかで使われる

Assumptions

1. The observed $y$ are independent, conditional on some predictors $x$
1. The response $y$ are normally distributed conditional on some predictors $x$
1. The response $y$ has constant variance, conditional on some predictors $x$
1. There is a straight line relationship between $y$ and the predictors $x$ and random effects $z$
1. Random effects $z$ are independent of $y$.
1. Random effects $z$ are normally distributed

```{r}
library(lme4)
```

```{r}
n <- 15
```

正規分布の乱数
```{r}
?rnorm()
```

* seed を設定して乱数の値を固定する
```{r}
set.seed(10)
rnorm(n, mean = 0, sd = 1)
```

```{r}
set.seed(10)
rnorm(n, mean = 0, sd = 1)
```
* data.frame の作成
```{r}
site_effect <- rnorm(n, mean=0, sd=1)
y1 <- site_effect + rnorm(n, mean=0, sd=1)
y2 <- site_effect + 6 + rnorm(n, mean=0, sd=1)

myDat <- data.frame(
  y = c(y1, y2), 
  treatment = c(rep("control", n), rep("fertil", n)), 
  site = rep(paste0("S", seq(1:n)), 2)
  )
str(myDat)
```
```{r}
head(myDat)
```

* mixed model
  * 固定効果: treatment
  * ランダム効果: site
```{r}
mod1 <- lmer(y ~ treatment + (1|site), data = myDat, REML=FALSE)
mod1
```
* Random effects の sd は 1 ぐらいになる
  * 乱数発生時のパラメータ
* treatmentfertil の Estimate: 6 ぐらいになる
  * データを作ったときの値
* p-value がない!

```{r}
summary(mod1)
```
```{r}
plot(mod1)
```

* p値の算出

```{r}
mod2<- lmer(y ~ 1 + (1|site), data = myDat, REML=FALSE)
mod2
```

```{r}
anova(mod1, mod2)
```

モデルの中身の取り出し

* `broom` package


```{r}
library(broom)
```


```{r}
tidy(mod1)
```
```{r}
library(dplyr)
```
```{r}
tidy(mod1) %>% 
  filter(term=="treatmentfertil") -> output
output$estimate
```

* 関数

```{r}
power_analysis <- function(n){
  site_effect <- rnorm(n, mean=0, sd=1)
  y1 <- site_effect + rnorm(n, mean=0, sd=1)
  y2 <- site_effect + 6 + rnorm(n, mean=0, sd=1)

  myDat <- data.frame(
    y = c(y1, y2), 
    treatment = c(rep("control", n), rep("fertil", n)), 
    site = rep(paste0("S", seq(1:n)), 2)
    )
  mod1 <- lmer(y ~ treatment + (1|site), data = myDat, REML=FALSE)
  tidy(mod1) %>% 
    filter(term=="treatmentfertil") -> output
  return(output$estimate)
}
power_analysis(5)
```
```{r}
power_analysis(2)
```

n=1 だとエラー
```{r}
power_analysis(1)
```

* 6に収束

```{r}
for (i in c(1:10)){
  n <- 2**i
  print(power_analysis(n))
}
```

```{r}
coef(mod1)
```

* 結果の違いをシミュレーションする

```{r}
data.frame(sample_size = rep(2:40, each=10)) %>% 
  rowwise() %>% 
  mutate(po = power_analysis(sample_size)) -> simulation
head(simulation)
str(simulation)
```

```{r}
library(ggplot2)
```

* 6に収束
```{r}
ggplot(simulation, aes(x = sample_size, y = po)) + 
  geom_point()
```

```{r}
compute_p <- function(n){
  site_effect <- rnorm(n, mean=0, sd=1)
  y1 <- site_effect + rnorm(n, mean=0, sd=1)
  y2 <- site_effect + 6 + rnorm(n, mean=0, sd=1)

  myDat <- data.frame(
    y = c(y1, y2), 
    treatment = c(rep("control", n), rep("fertil", n)), 
    site = rep(paste0("S", seq(1:n)), 2)
    )
  mod1 <- lmer(y ~ treatment + (1|site), data = myDat, REML=FALSE)
  mod2 <- lmer(y ~ 1 + (1|site), data = myDat, REML=FALSE)
  tst <- anova(mod1, mod2)
  tst %>% 
    tidy() %>% 
    filter(term=="mod1") -> output
  return(output[["p.value"]])
}
compute_p(5)
```

```{r}
suppressWarnings(
  data.frame(sample_size = rep(2:40, each=10)) %>% 
    rowwise() %>% 
    mutate(p.value = compute_p(sample_size)) -> simulation2
) 
```

```{r}
ggplot(simulation2, aes(x = sample_size, y = p.value)) + 
  geom_point()
```

```{r}
ggplot(simulation2, aes(x = sample_size, y = p.value)) + 
  geom_point() + 
  scale_y_continuous(trans = "log10")
```
